A fully labelled proof system for intuitionistic modal logics

### 2020-09-22 FullyLabelled/JLC/system.tex
moving remark to Sec 7

Now we can present the inference rules for \emph{system $\labIKp$} for the logic $\IK$.
We obtained this system, shown in Figure~\ref{fig:labIKp}, as follows.
%\sonia{This is an interesting alternative view but for me the starting point really is the Maffezioli, Naibo and Negri paper. Maybe rephrase slightly to acknowledge their work too.}
Our starting point was the multiple-conlusion nested sequent system \emph{\`a la} Maehara (as presented in~\cite{str:2017maehara}), which can straightforwardly translated into the labelled setting, and which yields the rules $\llabrn\bot$, $\llabrn\AND$, $\rlabrn\AND$, $\llabrn\OR$, $\rlabrn\OR$, $\llabrn\DIA$, and $\rlabrn\DIA$ as we show them in Figure~\ref{fig:labIKp}. However, this naive translation would also yield the rules $\rn{id'}$, $\llabrn\IMP'$, and~$\llabrn\BOX'$:
\begin{equation}
  \label{eq:nomon}
  \vlinf{\rn{id}'}{}{\B, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a} }{}
  \quad
  \vliinf{\llabrn\IMP'}{}{\B,\Left, \labels{x}{A \IMP B} \SEQ \Right}{\B,\labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left, \labels{x}{B} \SEQ \Right}
  \quad
  \vlinf{\llabrn\BOX}{}{\B, \accs xz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, \accs xz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}
\end{equation}
that are not sufficient for a complete system, and we will see below
why. Before, let us first look at the rules $\rlabrn\IMP$ and
$\rlabrn\BOX$. In the multiple-conlusion nested sequent system
of~\cite{str:2017maehara}, these are the two rules that force
single-conclusion. 
%This is the reason why in our system we need the
%$\le$ in the relational atoms; we need to go to the future of the
%current world.  
%
In our system, this phenomenon is replaced by a re-positioning of the considered formulas to a fresh label.
%
In the Kripke-semantics in~\eqref{eq:kripke} the two
connectives $\IMP$ and $\BOX$ are the ones that make use of the
pre-order relation $\le$. This relation is reflexive and transitive. In
order to capture that in the proof system, we need to add the rules
$\rn{refl}$ and $\rn{trans}$.\footnote{These are the same as for {\bf G3I} as presented in~\cite{dyckhoff2012}.}

%
%%These can be obtained by applying the axioms-as-rules methodology as in~\cite{maffezioli:etal:synthese13}.

Finally, in the semantics, the two relations $\rel$ and $\le$ are
strongly connected through the two conditions $\rn{F_1}$ and $\rn{F_2}$. These need to be reflected at the level of the proof system, which is done by the two rules $\rn{F_1}$ and $\rn{F_2}$.\footnote{The logic studied in~\cite{maffezioli2014} has similar semantic conditions (two binary relations in a Kripke frame) that lead to similar rules as ours in the proof system.}
%
However, these two rules create new labels, and in order to be
complete, the system needs the \emph{monotonicity} rule $\llabrn{mon}$, shown on the left below. 
\begin{equation}
  \label{eq:mono}
  \vlinf{\llabrn{mon}}{}{\lseq{\B, \futs{x}{y}}{\Left, \labels{x}{A}}\Right}{
    \lseq{\B, \futs{x}{y}}{\Left, \labels{x}{A}, \labels{y}{A}}\Right}
  \hskip6em
  \vlinf{\rlabrn{mon}}{}{\lseq{\B, \futs{x}{y}}{\Left}{\Right, \labels{y}{A}}}{
    \lseq{\B, \futs{x}{y}}{\Left}{\Right, \labels{x}{A}, \labels{y}{A}}}
\end{equation}
Since this rule is a form of contraction, it would cause the same
problems as contraction in a cut elimination proof. Hence, it is
preferable to have a system in which this rule is admissible. This is
the reason why we have monotonicity incorporated in the rules
$\rn{id}$, $\llabrn\IMP$ and $\llabrn\BOX$ in Figure~\ref{fig:labIKp},
instead of using the rules in~\eqref{eq:nomon}. Then, not only
$\llabrn{mon}$ but also its right-hand side version $\rlabrn{mon}$,
shown on the right in~\eqref{eq:mono} above become admissible.

@@

\todo{}
\sonia{
	The rule corresponding to the $F_1$ was obtained through the standard axioms-to-rule procedure that is well known for geometric axioms. 
%
A similar rule can already be found in~\cite{maffezioli2014}.
%
It corresponds to attributing a positive polarity to both relational and pre-order atoms.

$$\vlinf{}{\lb u \mbox{ fresh}}{\B, \accs xy, \futs yz, \Gamma \SEQ \Delta}{\accs xy, \futs yz, \futs xu, \accs uz, \Gamma \SEQ \Delta}$$

On the other hand, a different rule for $F_1$ could be obtained in the style of Vigano, that is, by polarising both relational and pre-order atoms negatively, as follows:

$$\vliiinf{}{\lb u \mbox{ fresh}}{\B, \Gamma \SEQ \Delta}{\B, \accs xy, \Gamma \SEQ \Delta}{\B, \futs yz, \Gamma \SEQ \Delta}{\futs xu, \accs uz, \Gamma \SEQ \Delta}$$

\lutz{we should say that this is pretty pointless wrt proof search.}
It is also possible to design a "mixed" rule, by considering relational atoms positively and pre-order atoms negatively:

$$\vliinf{}{\lb u \mbox{ fresh}}{\B, \accs xy, \Gamma \SEQ \Delta}{\B, \accs xy, \futs yz, \Gamma \SEQ \Delta}{\accs xy, \futs xu, \accs uz, \Gamma \SEQ \Delta}$$

Or vice-versa:

$$\vliinf{}{\lb u \mbox{ fresh}}{\B, \futs yz, \Gamma \SEQ \Delta}{\B, \futs yz, \accs xy, \Gamma \SEQ \Delta}{\futs yz, \futs xu, \accs uz, \Gamma \SEQ \Delta}$$
}

>>>>

Now we can present the inference rules for \emph{system $\labIKp$} for the logic $\IK$.
We obtained this system, shown in Figure~\ref{fig:labIKp}, as follows.
%\sonia{This is an interesting alternative view but for me the starting point really is the Maffezioli, Naibo and Negri paper. Maybe rephrase slightly to acknowledge their work too.}
Our starting point was the multiple-conlusion nested sequent system \emph{\`a la} Maehara (as presented in~\cite{str:2017maehara}), which can straightforwardly translated into the labelled setting, and which yields the rules $\llabrn\bot$, $\llabrn\AND$, $\rlabrn\AND$, $\llabrn\OR$, $\rlabrn\OR$, $\llabrn\DIA$, and $\rlabrn\DIA$ as we show them in Figure~\ref{fig:labIKp}. However, this naive translation would also yield the rules $\rn{id'}$, $\llabrn\IMP'$, and~$\llabrn\BOX'$:
\begin{equation}
  \label{eq:nomon}
  \vlinf{\rn{id}'}{}{\B, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a} }{}
  \quad
  \vliinf{\llabrn\IMP'}{}{\B,\Left, \labels{x}{A \IMP B} \SEQ \Right}{\B,\labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left, \labels{x}{B} \SEQ \Right}
  \quad
  \vlinf{\llabrn\BOX}{}{\B, \accs xz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, \accs xz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}
\end{equation}
that are not sufficient for a complete system, and we will see below
why. Before, let us first look at the rules $\rlabrn\IMP$ and
$\rlabrn\BOX$. In the multiple-conlusion nested sequent system
of~\cite{str:2017maehara}, these are the two rules that force
single-conclusion. 
%This is the reason why in our system we need the
%$\le$ in the relational atoms; we need to go to the future of the
%current world.  
%
In our system, this phenomenon is replaced by a re-positioning of the considered formulas to a fresh label.
%
In the Kripke-semantics in~\eqref{eq:kripke} the two
connectives $\IMP$ and $\BOX$ are the ones that make use of the
pre-order relation $\le$. This relation is reflexive and transitive. In
order to capture that in the proof system, we need to add the rules
$\rn{refl}$ and $\rn{trans}$.\footnote{These are the same as for {\bf G3I} as presented in~\cite{dyckhoff2012}.}
%
%%These can be obtained by applying the axioms-as-rules methodology as in~\cite{maffezioli:etal:synthese13}.

Finally, in the semantics, the two relations $\rel$ and $\le$ are
strongly connected through the two conditions $\rn{F_1}$ and $\rn{F_2}$. These need to be reflected at the level of the proof system, which is done by the two rules $\rn{F_1}$ and $\rn{F_2}$.\footnote{The logic studied in~\cite{maffezioli2014} has similar semantic conditions (two binary relations in a Kripke frame) that lead to similar rules as ours in the proof system.}
%
These rules require to create new labels, and in order to be
complete, the system needs the \emph{monotonicity} rule $\llabrn{mon}$, shown on the left below. 
\begin{equation}
  \label{eq:mono}
  \vlinf{\llabrn{mon}}{}{\lseq{\B, \futs{x}{y}}{\Left, \labels{x}{A}}\Right}{
    \lseq{\B, \futs{x}{y}}{\Left, \labels{x}{A}, \labels{y}{A}}\Right}
  \hskip6em
  \vlinf{\rlabrn{mon}}{}{\lseq{\B, \futs{x}{y}}{\Left}{\Right, \labels{y}{A}}}{
    \lseq{\B, \futs{x}{y}}{\Left}{\Right, \labels{x}{A}, \labels{y}{A}}}
\end{equation}
Since this rule is a form of contraction, it would cause the same
problems as contraction in a cut elimination proof. Hence, it is
preferable to have a system in which this rule is admissible. This is
the reason why we have monotonicity incorporated in the rules
$\rn{id}$, $\llabrn\IMP$ and $\llabrn\BOX$ in Figure~\ref{fig:labIKp},
instead of using the rules in~\eqref{eq:nomon}. Then, not only
$\llabrn{mon}$ but also its right-hand side version $\rlabrn{mon}$,
shown on the right in~\eqref{eq:mono} above become admissible.

@@
